# Using ZINB_NNGP on a real dataset

```{r}
library(mvtnorm)
library(Matrix)
library(reticulate)
library(lubridate)
library(sf)
library(ZINB.GP)
```


## Dataset
This time we will use some hurricane Ian damage data to show how to use the model on a sample real dataset. We start by loading some variables we are insterested in.

```{r}
responseData <- read.csv("examples/Ian/Ian Events Water.csv")

event_type <- model.matrix(~as.factor(responseData$EVENT_TYPE))
flood <- event_type[, 2]
tornado <- event_type[, 3]


startTime <- dmy_hms(responseData$BEGIN_DATE_TIME)
endTime <- dmy_hms(responseData$END_DATE_TIME)

# Make sure we don't accidentally have any identical times
timestamp <- (((endTime - startTime) / 2) + startTime) + rnorm(length(startTime), 0, 1)

time_distances <- as.matrix(dist(timestamp, diag = TRUE, upper = TRUE) / 100000)

places <- st_as_sf(responseData, coords = c("BEGIN_LON", "BEGIN_LAT"), crs = 3857)$geometry
coords <- cbind(responseData$BEGIN_LON, responseData$BEGIN_LAT)
coords <- coords[2:nrow(coords), ]
distances <- st_distance(places)

# Use direct deaths and highest nearby storm surge measurement as predictors
deaths <- responseData$DEATHS_DIRECT
max_water <- responseData$MAX_WATER

# Attempt to predict property damage
y <- responseData$DAMAGE_PROPERTY

Vs <- diag(length(y))[, 2:length(y)]
Ds <- distances
units(Ds) <- NULL
Ds <- Ds[2:nrow(Ds), 2:ncol(Ds)]

Vt <- diag(nrow(time_distances))[, 2:nrow(time_distances)]
Dt <- time_distances
Dt <- Dt[2:nrow(Dt), 2:ncol(Dt)]

X <- cbind(deaths, max_water)
X <- cbind(1, X)
```

## Running the model on the dataset
Now we are ready to run the model. This is done as follows:
```{r}
nsim <- 200
burn <- 100
M <- 10
thin <- 1

# Run for a short time for demo purposes
output <- ZINB_NNGP(X, y, coords, Vs, Vt, Ds, Dt, M, nsim, burn, thin, TRUE)
predictions <- output$Y_pred
sim_alpha <- output$Alpha
sim_beta <- output$Beta
```

Then we can investigate the output as desired, as an example, we create 90% CIs for alpha and beta, and investigate how often some samples were selected to be at risk.

Looking at 90% CIs for coefficients for the at-risk LR model:
```{r}
alpha
aCIs <- apply(sim_alpha, 2, function(x) quantile(x, probs=c(0.05, 0.95)))
boxplot(aCIs, main = "CIs for Alpha")
```

Looking at 90% CIs for coefficients for the NB model:
```{r}
beta
bCIs <- apply(sim_beta, 2, function(x) quantile(x, probs=c(0.05, 0.95)))
boxplot(bCIs, main = "CIs for Beta")
```

Viewing the frequency a few samples were at risk:
```{r}
# Examine how often various samples are at risk
at_risk <- output$at_risk
sim_p_at_risk <- apply(at_risk, 2, mean)
sim_p_at_risk[1:20]
```
